Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-06-03 vesicle_code/docs/permeableVes/main.tex
doing more shear runs
trying to finish shear example on new laptop
starting parabolic results
switching to PNAS

\documentclass[aps,prl,showpacs]{revtex4}

@@

%\usepackage{showkeys}
\usepackage{amsmath,amsfonts,bm}
\usepackage{graphicx,epsfig}
\usepackage{todonotes}
%Dfig puts a figure of specified size at the place indicated.
\newcommand{\Dfig}[2]{\epsfig{figure=#1,width=#2cm}}

%\Eonefigs puts one figure at the top of a page. 3 arguments.
%Usage:
%\Eonefigs{file fig}{caption}{size in inches}
%the file must be eps, but names must be given 
%WITHOUT .eps extension

\newenvironment{Eonefigs}[3]
{        \begin{figure} [t]
          \begin{center}
              \includegraphics[width=0.5\textwidth]{#3} \\
          \vspace{-0.21in}
          \caption{{\footnotesize #2}}
          \label{#1}
          \end{center}
        \vspace{-.3in}
        \end{figure}
}{}

\newcommand{\co}[1]{\marginpar{\scriptsize \textcolor{red}{#1}}} % red comment

@@

\title{Semi-permeable vesicles: mathematical formulation, numerics and physics}
\author{Ashley Gannon, Bryan Quaife, and Yuan-Nan Young}

\date{\today}

\begin{abstract}
We describe a mathematical model for a semi-permeable vesicle and derive
  a boundary integral equation formulation for simulating its
  hydrodynamics.  
\end{abstract}
\maketitle

\section{Problem formulation} \label{sc:formulate}
Consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid domain, subjected to an imposed flow
$\uu_\infty(\xx)$, for any $\xx \in \RR^2$.  Assume that the interior
and exterior fluids have the same viscosity $\mu$. Let $\xx$ be the
position of the interface $\gamma$, $\uu$ be the fluid velocity, and
$p$ be the pressure. In the vanishing Reynolds number limit, the
governing equations for the ambient fluid are
%
\begin{subequations}
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, 
  &&\xx \in \RR^2 \setminus \gamma, \\
% 
  \nabla \cdot \uu &= 0,  &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    \qquad &&\norm{\xx} \rightarrow \infty.
\end{alignat}
\end{subequations}%
%
Letting $\nn$ be the outward normal to $\gamma$, $\jump{\cdot}$ denote
the jump across the interface, and $T$ denote the hydrodynamic stress
tensor, the boundary conditions along the vesicle are
\begin{align}
  \label{eqn:jump1}
  \jump{T\cdot\nn} &= \ff_\mathrm{mem}, \\
  \label{eqn:jump2}
  \jump{\uu} &= 0,
\end{align}
where $\nn$ points outward to the vesicle, and $\ff_\mathrm{mem}$ is the
membrane force that includes a bending term that minimizes the Helfrich
energy, and a tension term that acts as a Lagrange multiplier to satisfy
the local inextensibility condition. The bending stiffness only sets the
time scale, so we let $k_b = 1$ throughout the paper. For ease of
implementation, we modify the bending tension and forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} + \ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -\xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
However, the physical tension $\Lambda$ can be computed as a
post-processing step as
\begin{align}
  \Lambda = \sigma + 1.5 \kappa^2,
\end{align}
where $\kappa$ is the membrane curvature.

The kinematic condition is modified to account for the fact that the
fluid can move through the membrane. A standard choice appears
is~\cite{}
\begin{align}
  \uu - \dot{\xx} = - \beta (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\beta$ is a semi-permeability rate.

Reduced area is $\nu = 4\pi A/L^2$.

\subsection{Integral equation formulation}
Since the flow is governed by the linear Stokes equations, it can be
represented in terms of layer potentials. Given the boundary
conditions~\eqref{eqn:jump1} and~\eqref{eqn:jump2} required on the
interface, the fluid velocity can be represented as the single-layer
potential
\begin{align}
  \uu(\xx) = \uu_\infty(\xx) + \SS[\ff_\mathrm{mem}](\xx), \quad
    \xx \in \Omega,
\end{align}
where
\begin{align}
  \SS[\ff](\xx) = \frac{1}{4\pi\mu} \int_{\gamma} \left(
    -\mathbf{I} \log\rho + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) ds_{\yy},
\end{align}
where $\rr = \xx - \yy$ and $\rho = \norm{\rr}$. Then, by the
kinematic equation, the dynamics of the vesicle 
\begin{align}
  \label{eqn:vesVelocity}
  \frac{d\xx}{dt} = \uu_\infty(\xx) + \beta (\ff_\mathrm{mem}\cdot\nn)\nn
  + \SS [\ff_\mathrm{mem}](\xx).
\end{align}
Finally, the inextensibility condition is given by
\begin{align}
  \xx_s\cdot\dot{\xx}_s = 0.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Area Dynamics}
Following the analysis of Veerapaneni et al.~\cite{vee-raj-bir-pur2009},
we compute the steady-state shape of a semi-permeable vesicle in a
quiescent flow. The steady-state shape is a stationary point of the
Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where $\Lambda$ is the constant tension that acts as a Lagrange
multiplier to satisfy the length conservation. Note that there is no
pressure Lagrange multiplier since the vesicle's area fluctuates. The
force acing on the membrane is found by taking the variation of
$\mathcal{L}$ with respect to $\gamma$, and by setting the total force
to zero, the steady state shape satisfies
\begin{align}
  \kappa_{ss} + \frac{1}{2}\kappa^3 - \Lambda \kappa = 0.
  \label{eqn:curvODE}
\end{align}
with initial conditions $\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.
Integrating the differential equation~\eqref{eqn:curvODE} with respect
to arclength, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \kappa_0^4 - \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curvODE2}
\end{align}
Assuming $\kappa_s \neq 0$, this differential equation is separable, and
can be solved analytically using Mathematica. For impermeable vesicles,
it is possible for the curvature to be
periodic~\cite{vee-raj-bir-pur2009}. However, the solution
of~\eqref{eqn:curvODE2} depends on the {\em EllipticF} function, and it
can not be made periodic. Therefore, the only solution
of~\eqref{eqn:curvODE2} is $\kappa(s) = \kappa_0$, meaning that the
steady state shape is a circle. By the inextensibility condition, the
steady state vesicle has a radius of $L/2\pi$, so its curvature and
tension are
\begin{align}
  \kappa = \frac{2\pi}{L}, \: \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:SSshape}
\end{align}
Note that in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

In addition to computing the steady shape, we can compute the transient
area of the vesicle. Parameterizing the vesicle with $\xx$, its area is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds .
  \label{eqn:area}
\end{align}
The derivative of $A(t)$ includes terms due to the time derivative of
$\xx \cdot \nn$, and a term due to interface
stretching~\cite{lai-tse-hua2008}. However, the stretching term is zero
because of the local inextensibility condition. Therefore,
\begin{align}
  \dot{A}(t) &= \frac{1}{2} \int_{\gamma} \pderiv{}{t} 
    (\xx \cdot \nn) \, ds = 
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
The second integrand can be computed as
\begin{align}
  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
\end{align}
Therefore,
\begin{align}
  \dot{A}(t) &= \int_\gamma (\dot{\xx} \cdot \nn)\, ds 
  = \int_\gamma \left(\beta (\ff \cdot \nn)\nn + \SS[\ff]\right) 
      \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality follows from the incompressibility of the
single-layer potential and the divergence theorem.  Using integration by
parts and the Frenet-Serret identities, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds.
\end{align}
Finally, using the relationship between the Lagrange multiplier
$\sigma$, and the tension $\Lambda$, the area of a semi-permeable
vesicle is governed by
\begin{align}
  \dot{A}(t) = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that this is consistent with the steady state curvature and
tension~\eqref{eqn:SSshape} since these constant values result in the
integrand in equation~\eqref{eqn:areaROC} being zero.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\subsection{Discretization in Space}
We apply a collocation method where the vesicle is discretized at $N$
points. We need to compute derivatives of functions defined on the
vesicle boundary, and these are done with Fourier differentiation.
Finally, to compute the single-layer potential, we use an eighth-order
Alpert quadrature rule~\cite{alp1999} which resolves the logarithmic
singularity of the integrand.

\subsection{Discretization in Time}
By using an integral equation formulation, the only unknowns are the
vesicle position and tension, and the flow in the bulk does not have to
be computed. However, the high-order derivatives in the bending and
tension forces~\eqref{eqn:membraneForces} result in stiffness if an
explicit time stepping method is applied to
equation~\eqref{eqn:vesVelocity}. We apply a similar semi-implicit time
stepping method to our previous works~\cite{vee-gue-zor-bir2009,
qua-bir2014}. First, given a vesicle $\gamma$, with a boundary
parameterized as $\xx(s)$, we introduce notation for all the required
differential and projection operators
\begin{align}
  B[\xx]\ff &= -\frac{d^4}{ds^4} \ff, \\
  T[\xx]\sigma &= (\sigma \xx_s)_s, \\
  D[\xx]\ff &= \xx_s \cdot \ff_s, \\
  P[\xx]\ff &= (\ff \cdot \nn) \nn. 
\end{align}
Here, the normal vector $\nn$ is the outward unit normal of $\gamma$,
and $d/ds$ is its arclength derivative. To simplify notation, we write
$B\ff$ to mean $B[\xx]\ff$, and use similar notation for the other
operators. Then, the dynamics of the vesicle are governed by
\begin{align}
  \frac{d\xx}{dt} &= \uu_{\infty}(\xx) + 
  \beta P(B\xx + T\sigma) + \SS(B\xx + T\sigma), \\
  D \frac{d\xx}{dt} &= 0.
\end{align}
By writing $B^N$ to mean the bending operator due to the vesicle
configuration at time $t^N$, and similar notation for the other
operators, we address the stiffness by introducing the semi-implicit
time stepping method
\begin{align}  
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \uu_\infty(\xx^N) 
  + \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) 
  + \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right).
\end{align}
Finally, the inextensibility condition is applied by requiring that 
\begin{align}
  D^N\xx^{N+1} = 1.
\end{align}

\todo[inline]{A few lines about SDC}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}
We consider a single semi-permeable vesicle in four different flow
regimes. All vesicles have a length of $L=4\pi$, but the area is not
conserved because of semi-permeability. However, the area is bounded by
$4\pi$ which is the area of a circle with length $L=4\pi$. For each
simulation, we apply an SDC second-order adaptive time stepping
method~\cite{qua-bir2016}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quiescent Flow} 
We consider a semi-permeable vesicle with initial reduced area $0.65$
and several different permeability rates $\beta$ in a quiescent flow. If
the vesicle is impermeable, meaning that $\beta = 0$, it will relax to a
steady-state shape with negative and constant
tension~\cite{kra-win-sei-lip1996}. However, from
equation~\eqref{eqn:SSshape}, a semipermeable vesicle relaxes to a
circle with positive constant tension. Since the area of a
semi-permeable vesicle is governed by~\eqref{eqn:areaROC}, we expect
that the time for the vesicle to deviate from its initial area, and the
time to reach the steady state will both scale with $\beta^{-1}$. In
Figure~\ref{fig:ellipseArea}, we plot the area of a vesicle with
different permeability rates. The area is computed directly from the
vesicle shape (solid lines) and by integrating
equation~\eqref{eqn:areaROC} (marks) with a first-order method. We also
include black dashed lines indicating the times, $t_{\mathrm{steady}}$,
that the vesicle is near its steady state circular shape. In
Table~\ref{tbl:ellipseRelaxTimes}, we report the time,
$t_\mathrm{steady}$, required for the vesicle's area to be within 1.3\%
of its maximum area, and the time, $t_\mathrm{inflate}$, required for
the vesicle's area to exceed 1.3\% of its initial area. We see that both
times scale with $\beta^{-1}$.

>>>>

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\title{Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and
Physics}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Ashley Gannon}
\author[b]{Yuan-Nan Young} 
\author[a,1]{Bryan Quaife}

\affil[a]{Department of Scientific Computing, Florida State University}
\affil[b]{Department of Mathematical Sciences, New Jeresey Institute of
Technology}

% Please give the surname of the lead author for the running footer
\leadauthor{Quaife}

% Please add a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement
about the significance of their research paper written at a level
understandable to an undergraduate educated scientist outside their
field of speciality. The primary goal of the significance statement is
to explain the relevance of the work in broad context to a broad
readership. The significance statement appears in the paper itself and
is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{All authors contributed equally.}
%\authordeclaration{Please declare any competing interests here.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
\correspondingauthor{\textsuperscript{1} E-mail: bquaife\@fsu.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
We describe a mathematical model for a semi-permeable vesicle and derive
  a boundary integral equation formulation for simulating its
  hydrodynamics.

Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

@@

%\usepgfplotslibrary{external}
%\tikzexternalize

@@

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{T}his PNAS journal template is provided to help you write your
work in the correct journal format. Instructions for use are provided
below. We do not start with a section heading for the introduction, but
rather type the intro starting here.

\section*{Governing equations} \label{sc:formulate}
Consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid domain, subjected to an imposed flow
$\uu_\infty(\xx)$, for any $\xx \in \RR^2$.  Assume that the interior
and exterior fluids have the same viscosity $\mu$. Let $\xx$ be the
position of the interface $\gamma$, $\uu$ be the fluid velocity, and
$p$ be the pressure. In the vanishing Reynolds number limit, the
governing equations for the ambient fluid are
%
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, 
  \nabla \cdot \uu = 0, \qquad &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    &&\norm{\xx} \rightarrow \infty,
\end{alignat}
where $\mu$ is the fluid viscosity. Letting $\nn$ be the outward normal
to $\gamma$, $\jump{\cdot}$ be the jump across the interface, and $T$ be
the hydrodynamic stress tensor, the interfacial boundary conditions are
\begin{align}
  \label{eqn:jump}
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad \jump{\uu} = 0,
\end{align}
where $\ff_\mathrm{mem}$ is the membrane force that results from the
Helfrich energy with bending modulus $k_b$, and a tension term that acts
as a Lagrange multiplier to satisfy the local inextensibility condition.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending tension and
forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} +
\ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -k_b \xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
The physical tension satisfies $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.

The kinematic condition is modified to account for flow through the
vesicle. A standard choice for this flux is the difference in mechanical
force~\cite{yao-mor2017}
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - \beta (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\beta$ is a semi-permeability rate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Integral equation formulation}
Since the flow is governed by the linear Stokes equations, it can be
represented in terms of layer potentials. Given the interfacial boundary
conditions~\eqref{eqn:jump}, the fluid velocity is
\begin{align}
  \uu(\xx) = \uu_\infty(\xx) + \SS[\ff_\mathrm{mem}](\xx), \quad
    \xx \in \Omega,
\end{align}
where $\SS$ is the single-layer potential
\begin{align}
  \SS[\ff](\xx) = \frac{1}{4\pi\mu} \int_{\gamma} \left(
    -\mathbf{I} \log\rho + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) ds_{\yy},
\end{align}
with $\rr = \xx - \yy$ and $\rho = \norm{\rr}$. Imposing the flux
condition~\eqref{eqn:vesicleFlux} and the inextensibility condition, the
vesicle velocity satisfies
\begin{align}
  \label{eqn:vesVelocity}
  \dot{\xx} &= \uu_\infty(\xx) + \beta (\ff_\mathrm{mem}\cdot\nn)\nn
  + \SS [\ff_\mathrm{mem}](\xx),  \quad
  \xx_s \cdot \dot{\xx}_s = 0.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Dimensionless equations}
The inextensibility condition guarantees that the vesicle's length, $L$,
is conserved, but its area, $A$, is no longer conserved because of
semi-permeability. Therefore, the reduced area $\nu = 4\pi A/L^2 \in
(0,1]$ is not conserved. We non-dimensionalize~\eqref{eqn:vesVelocity}
by introducing the length scale $R_0 = L/2\pi$ and defining the time
scale $\tau = \mu R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the
permeability scale $W = R_0/\mu$. Since the bending stiffness $k_b$ only
sets the time scale, we let $k_b=1$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Area dynamics}
Following the analysis of Veerapaneni et al.~\cite{vee-raj-bir-pur2009},
we compute the steady-state shape of a semi-permeable vesicle in a
quiescent flow. The steady-state shape is a stationary point of the
Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where the Lagrange multiplier $\Lambda$ the constant tension. Note that
there is no pressure Lagrange multiplier since the vesicle's area
fluctuates. Taking the variation of $\mathcal{L}$ with respect to
$\gamma$, and setting this vesicle force to zero, the steady state shape
satisfies $\kappa_{ss} + \frac{1}{2}\kappa^3 - \Lambda \kappa = 0$,
with initial conditions $\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.
Integrating once, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \kappa_0^4 - \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curvODE}
\end{align}
Assuming $\kappa_s \neq 0$, this differential equation is separable, and
can be solved analytically in terms of the {\em EllipticF} function
using Mathematica. Unlike in the impermeable
case~\cite{vee-raj-bir-pur2009}, the solution of~\eqref{eqn:curvODE} can
not be made periodic unless $\kappa(s) = \kappa_0$. Therefore, the
steady state shape of a semi-permeable vesicle in a quiescent flow is a
circle. By the inextensibility condition, the steady state vesicle has a
radius of $L/2\pi$, so its curvature and tension are
\begin{align}
  \kappa = \frac{2\pi}{L}, \: \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:SSshape}
\end{align}
Note that in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

In addition to computing the steady shape, we can compute the transient
area of the vesicle. The area of the vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
so the derivative of $A(t)$ includes terms due to the time derivative of
$\xx \cdot \nn$, and a term due to interface
stretching~\cite{lai-tse-hua2008}, but the stretching term is zero
because of the local inextensibility condition. Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts to the second integral, it is equivalent
to the first.
%\begin{align}
%  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
%    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
%    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
%    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
%    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
%\end{align}
Therefore,
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality follows from the incompressibility of the
single-layer potential.  Using integration by parts and the
Frenet-Serret identities, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds \\
  &= \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that this is consistent with the steady state curvature and tension
in~\eqref{eqn:SSshape} since these constant values result in the
integrand in~\eqref{eqn:areaROC} being zero.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Numerical Methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Space}
By using an integral equation formulation, the unknowns are the vesicle
position and tension which we track at a set of $N$ discretization
points. All arclength derivatives are computed with Fourier
differentiation, and eighth-order quadrature is applied to the weakly
singular single-layer potential~\cite{alp1999}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Time}
The high-order derivatives in the bending and tension
forces~\eqref{eqn:membraneForces} result in stiffness if an explicit
time stepping method is applied to equation~\eqref{eqn:vesVelocity}.
Therefore, we use a modification of a time stepping method that
discretizes high-order derivatives
semi-implicitly~\cite{vee-gue-zor-bir2009}. Introducing notation for
all the required differential and projection operators
\begin{alignat}{3}
  &B[\xx]\ff = -\frac{d^4}{ds^4} \ff,  \qquad
  &&T[\xx]\sigma = (\sigma \xx_s)_s, \\
  &D[\xx]\ff = \xx_s \cdot \ff_s, 
  &&P[\xx]\ff = (\ff \cdot \nn) \nn. 
\end{alignat}
the dynamics of the vesicle are governed by
\begin{align}
  \dot{\xx} = \uu_{\infty}(\xx) + 
  \beta P(B\xx + T\sigma) + \SS(B\xx + T\sigma), \quad
  D \dot{\xx} = 0.
\end{align}
Letting $B^N$ be the bending operator due to the vesicle
configuration at time $t^N$, and similar notation for the other
operators, we use the semi-implicit time stepping method
\begin{alignat}{3}  
  \frac{\xx^{N+1} - \xx^N}{\Delta t} &= \uu_\infty(\xx^N) 
  &&+ \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) \nonumber \\
  & &&+ \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right), \\
  D^N\xx^{N+1} &= 1.
\end{alignat}

To achieve high-order accuracy over long time horizons, we modify the
adaptive spectral deferred correction time stepping method introduced
in~\cite{qua-bir2016}. Since the area of the vesicle is no longer
conserved, we only use the length of the vesicle to estimate the
local truncation error which then determines if a time step is accepted
or rejected and the subsequent time step size.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Examples}
We consider a single semi-permeable vesicle in four different flow
regimes. All vesicles have a length of $L=4\pi$, but the area is not
conserved because of semi-permeability. However, the area is bounded by
$4\pi$ which is the area of a circle with length $L=4\pi$. For each
simulation, we apply an SDC second-order adaptive time stepping
method~\cite{qua-bir2016}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
We consider a semi-permeable vesicle with initial reduced area $\nu =
0.65$ and several different permeability rates $\beta$ in a quiescent
flow. If the vesicle is impermeable, meaning that $\beta = 0$, it will
relax to a steady-state shape with negative and constant
tension~\cite{kra-win-sei-lip1996}. However, from
equation~\eqref{eqn:SSshape}, a semi-permeable vesicle relaxes to a
circle with positive constant tension. Since the area of a
semi-permeable vesicle is governed by~\eqref{eqn:areaROC}, we expect
both the time for the vesicle to deviate from its initial area, and the
time to reach its steady state will scale with $\beta^{-1}$. In
Figure~\ref{fig:ellipseArea}, we plot the area of a vesicle with
different permeability rates. The area is computed directly from the
vesicle shape (solid lines) and by integrating
equation~\eqref{eqn:areaROC} (marks) with a first-order method. We also
include black dashed lines indicating the times, $t_{\mathrm{steady}}$,
that the vesicle is near its steady state circular shape. In
Table~\ref{tbl:ellipseRelaxTimes}, we report the time,
$t_\mathrm{steady}$, required for the vesicle's area to be within 1.3\%
of its maximum area, and the time, $t_\mathrm{inflate}$, required for
the vesicle's area to exceed 1.3\% of its initial area. We see that both
times scale with $\beta^{-1}$.

\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure}[htp]
  \centering
%  \ifTikz
%  \input{figures/ellipseArea.tikz}
%  \fi
  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The solid lines compute the area using
  equation~\eqref{eqn:area}, while the marks are computed by integrating
  equation~\eqref{eqn:areaROC}. The dashed black lines are the times
  when the vesicle's area is within 1.3\% of its maximum value of
  $4\pi$.}
\end{figure}
